### 2021-09-17 papers/modeling-of-combined-lfr-csp-sco2-paper/content.tex
finished the Results and Discussion section

The single cycle configuration in Wang et al. (2020) has a higher-temperature operating salt, 67\%KCl-33\%MgCl$_2$, as the heat transfer fluid in the CSP which is capable of temperatures in the range of 450 - 1400 $^{\circ}$C. The lower operating point of the studied solar salt in this paper, 60\%NaNO$_3$-40\%KNO$_3$ with a temperature range of 250 - 585 $^{\circ}$C, is employed because the technology is more established when compared to KCl-MgCl$_2$ \cite{turchi_2018}. The higher temperature salt requires preheating which is achieved prior to the CSP heat addition by a small modular lead-cooled fast reactor. The small modular LFR operates at a low temperature and a fixed location when compared to the higher temperature lead and variable location that the Westinghouse LFR studied in this paper is capable of. Due to the similar operating temperatures of the CSP and LFR, multiple cycle configurations are studied that are not possible with the similar components in Wang et al. (2020). Specifically studied in this paper are the effects on cycle and heat storage efficiency of different cycle configurations while having the LFR serve a dual purpose of electrical generation and a supplementary CSP TES charging technique. This paper will provide an overview of contending recompression sCO$_2$ Brayton cycles with varied positioning of complimentary CSP and LFR heat additions in the cycle. Additionally the location of where heat is drawn from the sCO$_2$ Brayton cycles to be stored in thermal energy storage is studied with the results discussed.

@@

Where $C_{avg}$ is the average capacitance rate across the hot or cold stream and $c_{in}$ and $c_{out}$ is the specific heat at the inlet and outlet respectively. Specific heat is found using library correlations. Once both average capacitances are calculated for the hot and cold streams, the one will have a larger value, $\dot{C}_{max}$, and one will have a smaller value,  $\dot{C}_{min}$. These maximum and minimum values are then used to find the capacitance ratio, $CR$, in Equation \ref{cap_ratio}.

@@

The C2S is located around the turbine and LFR at state points 5 to 6 depending on the charging cycle configuration: pre, parallel, post or circulator. In the C-1HTR1T-ON cycle, the LFR HX and CSP HX have identical inlet temperatures due to splitting the flow prior to their parallel orientation. Therefore, three sensitivity studies are done on the model. The initial two studies have the low LFR temperature constrained to the value of $400^{\circ}$C with varied cold CSP TES temperature and maximized cycle efficiency. To achieve a maximum cycle efficiency, the split fraction amount of flow to the main compressor, $y_{1}$, was parametrically studied.
%\mw{State what variable(s) are manipulated in order to maximize cycle efficiency -BW} 
Two cold TES temperatures are tested with constrained LFR low temperature of $400^{\circ}$C: 
\begin{itemize}
    \item	$410^{\circ}$C: Lowest cold TES temperature possible due to the sCO$_2$ cold inlet constrained from the LFR to $400^{\circ}$C and the addition of $10^{\circ}$C approach temperature;
    \item	$440^{\circ}$C: Upper bound temperature on cold TES storage;
\end{itemize}

@@

For all cycle configurations presented, with constrained or unconstrained LFR low temperature inlet and varied cold TES temperature, cycle and heat storage efficiencies were maximized using parametric studies on flow splitter fractions. These results were obtained using standardized values found in Table \ref{tab-cycle_sum} for a more direct comparison between cycles. Presentation and discussion of the results is fulfilled in the following section.

@@

C-CSP-ON is not affected by constrained or unconstrained LFR sCO$_2$ inlet temperature because the two cycles are not tethered during non-charging configuration. Two individual studies were conducted on the C-CSP-ON cycle cold TES temperature, one with a lower design value of 390$^{\circ}$C and another with a higher value of 440$^{\circ}$C. A cold TES temperature of 410$^{\circ}$C was excluded since it can be presumed to have an efficiency intermediate to the 390$^{\circ}$C and 440$^{\circ}$C calculated values. Cold TES temperature was found to have a negligible effect on cycle efficiency between the two studied cases, with the 440$^{\circ}$C case gaining 0.35\%. The component that has the largest change in values is the CSP HX with a large drop in capacitance ratio and UA value, therefore the lower temperature of 390$^{\circ}$C for cold TES temperature is ideal increasing the performances of the heat exchanger.

\subsubsection{C-1HTR1T-ON}
The single cycle configuration, C-1HTR1T-ON, has the CSP and LFR heat additions in parallel causing identical inlet conditions. Due to the 10$^{\circ}$C approach temperature, the identical inlet conditions lead to a lower bound of 410$^{\circ}$C on the cold TES temperature when the sCO$_2$ LFR inlet temperature is constrained to 400$^{\circ}$C. A supplementary 440$^{\circ}$C cold TES study is ran to test the effects that higher temperature storage has on cycle efficiency. Raising the temperature of the cold TES from 410$^{\circ}$C to 440$^{\circ}$C had no effect on cycle efficiency but did raise the solar salt mass flow rate to accommodate for the lower temperature difference across the CSP HX hot side. Increasing the solar salt mass flow rate requires more storage for the same CSP heat input, a reduction in dispatchability, and larger pumps, increasing the cost of the system. Therefore, a lower cold TES temperature is desirable. Without the constraint on the sCO$_2$ LFR inlet temperature, a study is additionally conducted with cold TES temperature set to 390$^{\circ}$C. Of these three sensitivity studies, the cycle that yielded the highest efficiency is the unconstrained, 390$^{\circ}$C cold TES with an efficiency gain of 0.68\% over the constrained cases. The high temperature difference between the hot and cold TES reduces the mass flow rate of solar salt and subsequent cost of system components. The sCO$_2$ LFR inlet temperature for the 390$^{\circ}$C cold TES study is 380$^{\circ}$C, which increases the performance of the LFR heat exchanger as well as allowing for a higher core power.

\subsubsection{C-2HTR3T-ON}
C-2HTR3T-ON has dedicated recuperators and turbines for the LFR and CSP, allowing for independent inlet temperatures. Three studies were conducted on this cycle with two being constrained and one unconstrained sCO$_2$ LFR inlet conditions. The constrained studies have cold TES temperatures of 390$^{\circ}$C and 440$^{\circ}$, both of which have negligible efficiency differences. Similarly to the studies done on C-1HTR1T-ON, the 440$^{\circ}$C cold TES temperature case has a larger mass flow rate of solar salt and therefore increases component cost. The 390$^{\circ}$C unconstrained study has the highest efficiency with a value of 46.10\%, 1.76\% more than the constrained cases. Additionally, the sCO$_2$ LFR inlet temperature is 415.2$^{\circ}$C, 15.2$^{\circ}$C more than constrained cases, causing heat exchanger parameters to be less than ideal but increasing LFR efficiency.

@@

\subsubsection{C-LFR-PRE}
The TES charging technique with the turbine prior to the C2S heat exchanger, C-LFR-PRE, requires the temperature of the C2S inlet to be high enough to charge the hot TES. The hot TES is at a temperature of 560$^{\circ}C$, with a 10$^{\circ}C$ approach temperature, therefore the inlet temperature to C2S is required to be at least 570$^{\circ}C$. In cases where the turbine outlet temperature is less than 570$^{\circ}C$, a isenthalpic valve bypasses flow from the high temperature turbine inlet and mixes with the low temperature turbine outlet raising the temperature prior to the C2S heat exchanger. For the C-LFR-PRE configuration to be viable, it requires a larger temperature than the set LFR HX outlet temperature of 595$^{\circ}C$. With this low of a LFR outlet temperature, 73.78\% of the flow is redirected around the turbine to raise the temperature prior to C2S. As the temperature of the hot TES is raised above a value of 540$^{\circ}$, the amount of flow around the turbine approaches 100\% and balancing the turbine power to compressor power is infeasible. The temperature in the hot TES is therefore set to a lower value of 540$^{\circ}$ to allow for operation of this cycle. With heat storage efficiency maximized, 93.89\% of the flow is sent to the main compressor, rejecting 622 MW of heat out of the cycle through the precooler. This amount of rejected heat is reflected in a heat storage efficiency of 34.53\%, the lowest of all tested TES charging configurations.

@@

To test the effects of splitting flow prior to the turbine, a parallel configuration C-LFR-PAR is studied. To reduce the pressure from 28.8 MPa to 8.8 MPa, an isenthalpic valve is required after heat is extracted through C2S.

@@

>>>>

The single cycle configuration in Wang et al. (2020) has a higher-temperature operating salt, 67\%KCl-33\%MgCl$_2$, as the heat transfer fluid in the CSP which is capable of temperatures in the range of 450 - 1400 $^{\circ}$C. The lower operating point of the studied solar salt in this paper, 60\%NaNO$_3$-40\%KNO$_3$ with a temperature range of 250 - 585 $^{\circ}$C, is employed because the technology is more established when compared to KCl-MgCl$_2$ \cite{turchi_2018}. The higher temperature salt requires preheating which is achieved prior to the CSP heat addition by a small modular lead-cooled fast reactor. The small modular LFR operates at a low temperature and a fixed location when compared to the higher temperature lead and variable location that the Westinghouse LFR studied in this paper is capable of. Due to the similar operating temperatures of the CSP and LFR, multiple cycle configurations are studied that are not possible with the similar components in Wang et al. (2020). Specifically studied in this paper are the effects on cycle and heat storage efficiency of different cycle configurations while having the LFR serve a dual purpose of electrical generation and a supplementary CSP TES charging technique. This paper provides an overview of contending recompression sCO$_2$ Brayton cycles with varied positioning of complimentary CSP and LFR heat additions in the cycle. Additionally the location of where heat is drawn from the sCO$_2$ Brayton cycles to be stored in thermal energy storage is studied with the results discussed.

@@

Where $C_{avg}$ is the average capacitance rate across the hot or cold stream and $c_{in}$ and $c_{out}$ is the specific heat at the inlet and outlet respectively. Specific heat is found using library correlations. Once both average capacitances are calculated for the hot and cold streams, one has a larger value, $\dot{C}_{max}$, and one has a smaller value,  $\dot{C}_{min}$. These maximum and minimum values are used to find the capacitance ratio, $CR$, in Equation \ref{cap_ratio}.

@@

The C2S is located around the turbine and LFR at state points 5 to 6 depending on the charging cycle configuration: pre, parallel, post or circulator. In the C-1HTR1T-ON cycle, the LFR HX and CSP HX have identical inlet temperatures due to splitting the flow prior to their parallel orientation. Therefore, three sensitivity studies are done on the model. The initial two studies have the low LFR temperature constrained to the value of $400^{\circ}$C with varied cold CSP TES temperature and maximized cycle efficiency. To achieve a maximum cycle efficiency, the split fraction amount of flow to the main compressor, $y_{1}$, is parametrically studied.
%\mw{State what variable(s) are manipulated in order to maximize cycle efficiency -BW} 
Two cold TES temperatures are tested with constrained LFR low temperature of $400^{\circ}$C: 
\begin{itemize}
    \item	$410^{\circ}$C: Lowest cold TES temperature possible due to the sCO$_2$ cold inlet constrained from the LFR to $400^{\circ}$C and the addition of $10^{\circ}$C approach temperature;
    \item	$440^{\circ}$C: Upper bound temperature on cold TES storage;
\end{itemize}

@@

For all cycle configurations presented, with constrained or unconstrained LFR low temperature inlet and varied cold TES temperature, cycle and heat storage efficiencies are maximized using parametric studies on flow splitter fractions. These results are obtained using standardized values found in Table \ref{tab-cycle_sum} for a more direct comparison between cycles. Presentation and discussion of the results is fulfilled in the following section.

@@

C-CSP-ON is not affected by constrained or unconstrained LFR sCO$_2$ inlet temperature because the two cycles are not tethered during non-charging configuration. Two individual studies are conducted on the C-CSP-ON cycle cold TES temperature, one with a lower design value of 390$^{\circ}$C and another with a higher value of 440$^{\circ}$C. A cold TES temperature of 410$^{\circ}$C is excluded since it can be presumed to have an efficiency intermediate to the 390$^{\circ}$C and 440$^{\circ}$C calculated values. Cold TES temperature is found to have a negligible effect on cycle efficiency between the two studied cases, with the 440$^{\circ}$C case gaining 0.35\%. The component that has the largest change in values is the CSP HX with a large drop in capacitance ratio and UA value, therefore the lower temperature of 390$^{\circ}$C for cold TES temperature is ideal increasing the performances of the heat exchanger.

\subsubsection{C-1HTR1T-ON}
The single cycle configuration, C-1HTR1T-ON, has the CSP and LFR heat additions in parallel causing identical inlet conditions. Due to the 10$^{\circ}$C approach temperature, the identical inlet conditions lead to a lower bound of 410$^{\circ}$C on the cold TES temperature when the sCO$_2$ LFR inlet temperature is constrained to 400$^{\circ}$C. A supplementary 440$^{\circ}$C cold TES study is ran to test the effects that higher temperature storage has on cycle efficiency. Raising the temperature of the cold TES from 410$^{\circ}$C to 440$^{\circ}$C has no effect on cycle efficiency but did raise the solar salt mass flow rate to accommodate for the lower temperature difference across the CSP HX hot side. Increasing the solar salt mass flow rate requires more storage for the same CSP heat input, a reduction in dispatchability, and larger pumps, increasing the cost of the system. Therefore, a lower cold TES temperature is desirable. Without the constraint on the sCO$_2$ LFR inlet temperature, a study is additionally conducted with cold TES temperature set to 390$^{\circ}$C. Of these three sensitivity studies, the cycle that yielded the highest efficiency is the unconstrained, 390$^{\circ}$C cold TES with an efficiency gain of 0.68\% over the constrained cases. The high temperature difference between the hot and cold TES reduces the mass flow rate of solar salt and subsequent cost of system components. The sCO$_2$ LFR inlet temperature for the 390$^{\circ}$C cold TES study is 380$^{\circ}$C, which increases the performance of the LFR heat exchanger as well as allowing for a higher core power.

\subsubsection{C-2HTR3T-ON}
C-2HTR3T-ON has dedicated recuperators and turbines for the LFR and CSP, allowing for independent inlet temperatures. Three studies are conducted on this cycle with two being constrained and one unconstrained sCO$_2$ LFR inlet conditions. The constrained studies have cold TES temperatures of 390$^{\circ}$C and 440$^{\circ}$, both of which have negligible efficiency differences. Similarly to the studies done on C-1HTR1T-ON, the 440$^{\circ}$C cold TES temperature case has a larger mass flow rate of solar salt and therefore increases component cost. The 390$^{\circ}$C unconstrained study has the highest efficiency with a value of 46.10\%, 1.76\% more than the constrained cases. Additionally, the sCO$_2$ LFR inlet temperature is 415.2$^{\circ}$C, 15.2$^{\circ}$C more than constrained cases, causing heat exchanger parameters to be less than ideal but increasing LFR efficiency.

@@

\subsubsection{C-LFR-PRE}
The TES charging technique with the turbine prior to the C2S heat exchanger, C-LFR-PRE, requires the temperature of the C2S inlet to be high enough to charge the hot TES. The hot TES is at a temperature of 560$^{\circ}$C, with a 10$^{\circ}$C approach temperature, therefore the inlet temperature to C2S is required to be at least 570$^{\circ}$C. In cases where the turbine outlet temperature is less than 570$^{\circ}$C, a isenthalpic valve bypasses flow from the high temperature turbine inlet and mixes with the low temperature turbine outlet raising the temperature prior to the C2S heat exchanger. For the C-LFR-PRE configuration to be viable, it requires a larger temperature than the set LFR HX outlet temperature of 595$^{\circ}$C. With this low of a LFR outlet temperature, 73.78\% of the flow is redirected around the turbine to raise the temperature prior to C2S. As the temperature of the hot TES is raised above a value of 540$^{\circ}$C, the amount of flow around the turbine approaches 100\% and balancing the turbine power to compressor power is infeasible. The temperature in the hot TES is therefore set to a lower value of 540$^{\circ}$C to allow for operation of this cycle. With heat storage efficiency maximized, 93.89\% of the flow is sent to the main compressor, rejecting 622 MW of heat out of the cycle through the precooler. This amount of rejected heat is reflected in a heat storage efficiency of 34.53\%, the lowest of all tested TES charging configurations.

@@

To test the effects of splitting flow prior to the turbine, a parallel configuration C-LFR-PAR is studied. To reduce the pressure from 28.8 MPa to 8.8 MPa, an isenthalpic valve is required after heat is extracted for storage through C2S. With this configuration two different values for cold TES are studied, one with a cold TES storage of 390$^{\circ}$C the other with a temperature of 440$^{\circ}$C. A noteworthy difference for this cycle is the C2S approach temperature is not set to 10$^{\circ}$C. Defining the approach temperature as 10$^{\circ}$C causes the energy balance governing combiner 2 to be over constrained. Therefore this value is free and calculated to be 35$^{\circ}$C for the 390$^{\circ}$C case and 26.27$^{\circ}$C for the 440$^{\circ}$C case. Both of these different temperature cold TES studies do not change the cycle parameters because, as seen in C-1HTR1T-ON and C-2HTR3T-ON, the solar salt mass flow rate adjusts to accommodate for the temperature difference. The higher cold TES temperature of 440$^{\circ}$C causes the mass flow rate in the CSP to increase by approximately 650 kg/s. Increasing the mass flow rate has a higher demand on system components but in this situation it allows for the salt to be charged at a higher rate, increasing the storage capabilities. Both cycles have a heat storage efficiency of 45.30\%. \hl{Should I include solar salt mass flow also in table 4?}

@@

C-LFR-CIRC is a thermal energy storage charging technique which does not suffer from losses associated with turbines, compressors, and heat exchangers. This is due to the fact that a separate recirculation cycle is modeled taking heat directly from the LFR, passing it through a recirculating pump, exchanges the heat into the hot TES through C2S, then removes heat in a chiller to bring the temperature of the flow down to the inlet temperature of the LFR, 400$^{\circ}$C. This is the most adoptable TES charging technique a wide range of possible TES charging temperatures with almost all of the heat from the LFR being transferred into the hot TES depending on the temperature. Multiple temperatures for the cold TES are studied; 390$^{\circ}$C, 410$^{\circ}$C, and 440$^{\circ}$C. Of these three temperatures, the highest efficiency is a cold TES temperature of 390$^{\circ}$ with a heat storage efficiency of 99.92\%. The heat storage efficiency reduces as the temperature of the cold TES is increased, the 410$^{\circ}$C case has an efficiency of 89.66\% while the 440$^{\circ}$C case has an efficiency of 74.29\%. This reduction in efficiency is due to the inlet temperature condition on the LFR, the higher the temperature of the cold TES the higher the temperature of the C2S outlet and therefore the more heat is being extracted through the chiller to bring the flow temperature down to the LFR inlet temperature. A drawback of this TES charging technique is the additional components needed for the circulation loop, with an additional chiller and circulator component, adding cost.
